---
title: "Hockey Teams in Pennsylvania"
author: "Macy Trout"
date: "2024-12-01"
output:
  html_document:
    code_folding: hide
    theme: paper
    highlight: kate
---

This project maps the locations and arena capacities of Pennsylvania’s professional hockey teams, including the NHL’s Flyers and Penguins and the AHL’s Phantoms, Bears, and Wilkes-Barre/Scranton Penguins. The design incorporates hockey-inspired elements like icy blue tones, star-shaped markers, and a minimalist layout to reflect the sport’s theme.

Map Description and Design Considerations

Map 1: Hockey Team Locations

This map displays team locations, with star sizes representing arena capacities. The dark blue background and white/blue stars evoke a hockey rink, while clean labels ensure clarity.

Key Design Choices:
- Color Palette: Dark blue tones for the background; white and light blue for stars to emphasize the icy theme.

- Point Representation: Stars sized by arena capacity to reflect audience potential.

- Labels: Simple and clear, positioned to avoid crowding.

Reflection

This map effectively visualizes Pennsylvania’s hockey teams, balancing aesthetics and analysis. Arena capacities provide context beyond geography, while the clean design and cool tones align with hockey’s visual identity.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

teams <- data.frame(
  Team = c("Philadelphia Flyers", "Pittsburgh Penguins", 
           "Wilkes-Barre/Scranton Penguins", "Lehigh Valley Phantoms"),
  League = c("NHL", "NHL", "AHL", "AHL"),
  City = c("Philadelphia", "Pittsburgh", "Wilkes-Barre", "Allentown"),
  Longitude = c(-75.172, -80.005, -75.880, -75.490),
  Latitude = c(39.950, 40.440, 41.245, 40.611)
)


teams_sf <- st_as_sf(teams, coords = c("Longitude", "Latitude"), crs = 4326)

```




```{r, warning=FALSE, message=FALSE}
library(devtools)
library(rnaturalearthhires)
library(rnaturalearth)
library(sf)


usa <- ne_states(country = "United States of America", returnclass = "sf")


library(dplyr)


pennsylvania <- usa %>% filter(name == "Pennsylvania")
```


```{r hockey-map, echo=TRUE, fig.width=8, fig.height=6, dpi=150, fig.align="center"}
library(ggplot2)
library(sf)
library(showtext)

font_add_google("Libre Baskerville", "libre_baskerville")
showtext_auto()

teams <- data.frame(
  Team = c("Philadelphia Flyers", "Pittsburgh Penguins", 
           "Wilkes-Barre/Scranton Penguins", "Lehigh Valley Phantoms", "Hershey Bears"),
  League = c("NHL", "NHL", "AHL", "AHL", "AHL"),
  Longitude = c(-75.172, -80.005, -75.880, -75.490, -76.652),
  Latitude = c(39.950, 40.440, 41.245, 40.611, 40.285),
  Capacity = c(19500, 18387, 8300, 8500, 10500)
)

ggplot() +
  geom_sf(data = pennsylvania, fill = "#202A44", color = "white", size = 0.8) +
  geom_point(data = teams, aes(x = Longitude, y = Latitude, size = Capacity), 
             shape = 8, fill = "white", color = "#ADD8E6", stroke = 1.2) +
  geom_text(data = teams, aes(x = Longitude, y = Latitude, label = Team),
            vjust = -1.5, size = 4, family = "libre_baskerville", color = "white") +
  scale_size_continuous(range = c(4, 10), breaks = c(8000, 10000, 15000, 20000),
                        name = "Arena Capacity") +
  coord_sf(xlim = c(-81, -74), ylim = c(39, 42.5), expand = FALSE) +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#1B2735", color = NA),
    panel.background = element_blank(),
    plot.title = element_text(size = 22, face = "bold", hjust = 0.5, family = "libre_baskerville", color = "#ADD8E6"),
    plot.subtitle = element_text(size = 16, hjust = 0.5, family = "libre_baskerville", color = "white"),
    legend.position = "bottom",
    legend.title = element_text(size = 14, face = "bold", family = "libre_baskerville", color = "white"),
    legend.text = element_text(size = 12, family = "libre_baskerville", color = "white"),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_line(color = "#2E4C6D", size = 0.2)
  ) +
  labs(
    title = "Hockey Teams in Pennsylvania",
    subtitle = "NHL and AHL Teams with Arena Capacities",
    size = "Arena Capacity"
  ) 
```



